---
title: "DATA 607 Week 9 Assignment"
author: "Ahsanul Choudhury"
date: "October 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Objective:
To choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.
***

For my assignment I chose to work with the New York Times "Most Popular API". Under "Most Popular API" there are three sections: 

-  Most Emailed by Section & Time Period

-  Most Shared by Section & Time Period

-  Most Viewed by Section & Time Period
I am interested in using the *Most Viewed* section to see which Titles are most viewed today. New York Times "Most Popular API" give 20 results per page so I will use the top 20 results and plot the numeber of times each sections appears in top 20 most viewed today and for the past 7 days to get an idea of readers interest.

####API Key
The first step for this assignment was to register and get a API key from New York Times. Once the API key was received the following steps were taken to complete the assignment

####Load Libraries
```{r library, include=TRUE}
library(jsonlite)
library(plyr)
library(dplyr)
library(knitr)
library(ggplot2)
```

####Establish connection with New York Times
```{r load_data, include=TRUE}
uri <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/mostviewed/all-sections/1.json?api-key=31f8d708d3d9492c906146c5a1215690")
```


####Create R Data Frame
```{r create_data_frame, include=TRUE}
df <- ldply(uri[4], data.frame)
```


####Get details of data frame
```{r get_details, include=TRUE}
length(df)
names(df)
```


####Edit data frame
```{r select_column, include=TRUE}
MostViewed <- df %>% select(views,title, section, url, byline, type)
head(MostViewed)
```

####Final data frame
```{r create_table, include=TRUE}
kable(MostViewed)
```


#####I wanted to see how many times each section appears in the top 20 most viewed today and in past 7 days

####Section count in top 20 most viewed today
```{r section_count, include=TRUE}
SectionCount <- MostViewed %>%count(section)
SectionCount
```


####Plot section count in top 20 most viewed today
```{r plot_today, include=TRUE}
ggplot(data = SectionCount, aes(x = section, y = n, fill = section))+ geom_bar(stat="identity", position="dodge") + ggtitle("Most Viewed by Section (Today)") + ylab("Frequency")+ coord_flip() + geom_text(aes(label = n))
```

####Plot section count in top 20 most viewed for the past 7 days
```{r plot_week, include=TRUE}
uriWeek <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/mostviewed/all-sections/7.json?api-key=31f8d708d3d9492c906146c5a1215690")
dfWeek <- ldply(uriWeek[4], data.frame)
SectionCountWeek <- dfWeek %>%count(section)
ggplot(data = SectionCountWeek, aes(x = section, y = n, fill = section))+ geom_bar(stat="identity", position="dodge") + ggtitle("Most Viewed by Section(1 Week)") + ylab("Frequency")+ coord_flip() + geom_text(aes(label = n))
```

###Conclusion
With API New York time provide an easy way to access their data. Once I was able to connect using API the JSON data is easy to manipulate with in r and there are few libraries available for that.