---
title: "607 Week3"
author: "Ahsanul Choudhury"
date: "September 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#required Library
```{r library, include=TRUE}
library(stringr)
```


##3. Copy the introductory example. The vector name stores the extracted names.

##Load Data
```{r load_data, include=TRUE}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```


##(a)Use the tools of this chapter to rearrnge the vector so that all elements conform to the standard first_name last_name'

##Remove Title
```{r Remove1, include=TRUE}
noTitle <- unlist(str_replace(name,"[[:alpha:]]{2,}[.]",""))
noTitle
```

##Remove Second Name
```{r remove2, include=TRUE}
noSecName <- unlist(str_replace(noTitle,"[[:alpha:]]{1}[.]",""))
noSecName
```

##Extracting First Name
```{r First_name, include=TRUE}
firstName <- unlist(str_extract(noSecName,"[[:alpha:]]{2,}[[:space:]]{1,}|[[:punct:]][[:space:]]{1,}[[:alpha:]]{2,}"))
firstName <- unlist(str_replace(firstName,"[[:punct:]][[:space:]]",""))
firstName
```


#Extracting Last Name
```{r last_name, include=TRUE}
lastName <- unlist(str_extract(noSecName,"[^[:punct:]][[:space:]][[:alpha:]]{2,}|[[:alpha:]]{2,}[[:punct:]]"))
lastName <- unlist(str_replace(lastName,"[[:alpha:]][[:space:]]", ""))
lastName <- unlist(str_replace(lastName,"[[:punct:]]", ""))
lastName
```
##For first names and last names I was having trobule with mixing characters and punctuations, used str_replace function to clean up the data.

##Rearrange
```{r rearrange, include=TRUE}
paste(firstName, lastName)
data.frame(firstName, lastName)
```


##(b)Construct a logical vector indicating whether a character has a title(i.e., Rev. and Dr.).

##Title has more than 2 alphabet and ends with"."
```{r title, include=TRUE}
title <- unlist(str_detect(name,"[[:alpha:]]{2,}[.]"))
title
df1 <- data.frame(name,title)
df1
```


##(c)Construct a logical vector indicating whether a character has a second name.

##Second Name
```{r 2ndName, include=TRUE}
secName <- unlist(str_detect(noTitle,"[[:alpha:]]{1,}[.]"))
secName
df2 <- data.frame(name,secName)
df2
```



##Describe the types of strings that conform to the following regular expressions and construct and example that is matched by regular expression.

##(a)[0-9]+\\$

##The above regular expression looking for numbers 0 to and ends with $.
```{r sol1, include=TRUE}
sol1 <- c("124$", "$124", "a2b$", "16$cb")
sol1 <- unlist(str_detect(sol1, "[0-9]+\\$"))
sol1
```

##(b)\\b[a-z]{1,4}\\b

##This regular expression will extract data containing blank followed by 1 to four lower case letters followed by blank.
```{r sol2, include=TRUE}
sol2 <- c("abcdf", "ghij", "KLMN", "6542", "opqr 23")
sol2 <- unlist(str_extract(sol2, "\\b[a-z]{1,4}\\b"))
sol2
```

##(c).*?\\.txt$
##This regular expression will look for files ends in .txt
```{r sol3, include=TRUE}
sol3 <- c("ahs.rtxt", "123asd.txt", "txt.r")
sol3 <- unlist(str_extract(sol3, ".*?\\.txt$"))
sol3
```

##(d)\\d{2}/\\d{2}\\d{4}
##This regular expression is looking for two numbers followed by a forward slash followed by two numbers followed by a forward slash followed by 4 numbers, most likely a date format.
```{r sol4, include=TRUE}
sol4 <- c("09/24/1979", "mm/dd/yyyy", "23/sept/1979")
sol4 <- unlist(str_extract(sol4, "\\d{2}/\\d{2}/\\d{4}"))
sol4
```

##(e)<(.+?)>.+?</\\1>
##This regular expression will look for html tags
```{r sol5, include=TRUE}
sol5 <- c("<tag>cuny msda</tag>", "<tag>Assingnment2<tag>", "www.cuny.com")
sol5 <- unlist(str_extract(sol5, "<(.+?)>.+?</\\1>"))
sol5
```



##9.The following code hides a secret message. Crack it with R and regular expressions.

##Load Hidden message
```{r hMessage, include=TRUE}
hMessage <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
```
##Upper cases are more revealing than the others and "." used as word sepatators


##Extract characters
```{r characters, include=TRUE}
message <- unlist(str_extract_all(hMessage, "[[:upper:].]{1,}"))
message
```

##Join and Replace "." with space
```{r message, include=TRUE}
message <- str_replace_all(paste(message, collapse = ''), "[.]", " ")
message
```